---
title: "Homework 2"
author: "PSTAT 131/231"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Linear Regression

```{r}
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
library(yardstick)
tidymodels_prefer()

```

### Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.

The age histogram is right skewed means that more of the abalone has age less than 15.
```{r}

abalone <- read.csv(file = 'abalone.csv')
abalone['age']=abalone['rings']+1.5

ggplot(abalone, aes(x=age)) + geom_histogram(binwidth=1)


```

### Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.
```{r}
set.seed(3435)

abalone_split <- initial_split(abalone, prop = 0.80,
                                strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)


a <- c(16.5,8.5,10.5,11.5,8.5,9.5,21.5,17.5,10.5,20.5)


b <- scale(a,center=TRUE,scale=TRUE)
```
*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

### Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.


```{r}
# we are not include 'rings' to predict age because age is directly calculated from rings


abalone_recipe <- recipe(age ~ ., data = abalone_train) %>% 
  step_rm('rings')%>%
  step_dummy(all_nominal_predictors())%>%
  step_interact(terms = ~type:shucked_weight)%>%
  step_interact(terms = ~longest_shell:diameter)%>%
  step_interact(terms = ~shucked_weight:shell_weight)%>%
  step_center(all_predictors())%>%
  step_scale(all_predictors())

```
### Question 4

Create and store a linear regression object using the `"lm"` engine.

```{r}

lm_model <- linear_reg() %>% 
  set_engine("lm")


```



### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.

```{r}
myWorkflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)



```





### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
lm_fit <- fit(myWorkflow,abalone_train)

lm_fit %>% 
  # This returns the parsnip object:
  extract_fit_parsnip() %>% 
  # Now tidy the linear model object:
  tidy()

predicted_age <- 11.42335329+0.5*0.85928463+0.1*2.37253150+0.3*0.25076110+
  4*4.24434255+1*-3.65543589+2*-0.81318578+1*1.88006719-0.33199999+0.02721407-
  3.26447671-0.22901120
  

```



### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).
2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).
3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.


```{r}

abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-c(age)))

abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))

  

abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics(abalone_train_res, truth = age, 
                estimate = .pred)


# In our model,r-squared is 0.5464 reveals that 54% of the data fit the regression model.
```



### Required for 231 Students

In lecture, we presented the general bias-variance tradeoff, which takes the form:

$$
E[(y_0 - \hat{f}(x_0))^2]=Var(\hat{f}(x_0))+[Bias(\hat{f}(x_0))]^2+Var(\epsilon)
$$

where the underlying model $Y=f(X)+\epsilon$ satisfies the following:

- $\epsilon$ is a zero-mean random noise term and $X$ is non-random (all randomness in $Y$ comes from $\epsilon$);
- $(x_0, y_0)$ represents a test observation, independent of the training set, drawn from the same model;
- $\hat{f}(.)$ is the estimate of $f$ obtained from the training set.

#### Question 8

Which term(s) in the bias-variance tradeoff above represent the reproducible error? Which term(s) represent the irreducible error?

#### Question 9

Using the bias-variance tradeoff above, demonstrate that the expected test error is always at least as large as the irreducible error.

#### Question 10

Prove the bias-variance tradeoff.

Hints:

- use the definition of $Bias(\hat{f}(x_0))=E[\hat{f}(x_0)]-f(x_0)$;
- reorganize terms in the expected test error by adding and subtracting $E[\hat{f}(x_0)]$